Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems

### 2020-10-01 2020/OPTIMA/slides/slides.tex
OPTIMA: final update of slides

\begin{frame}
  \frametitle{Problem statement}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{displaymath}
        \begin{array}{cr}\\
          \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y:\\
           g_j(y)\leqslant 0, 1\leqslant j\leqslant m\},
        \end{array}
      \end{displaymath}
      where \(D\) is a hyperrectangle,
      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
      \begin{displaymath}
        \begin{array}{cr}\\
          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
        \end{array}
      \end{displaymath}
      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Problem statement}
  Such sets of problems can arise in the following cases:
  \begin{itemize}
    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
    \item a set of problems, obtained after scalarization of a multi-objective problem.
  \end{itemize}
\textbf{Possible approaches}:
  \begin{itemize}
    \item solve each problem independently;
    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
    prioritizing one of the problems.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Algorithm for solving a set of problems}
  The method relies on the previously introduces AGS and consists of the following steps:
  \begin{itemize}
    \item Create \(q\) copies of AGS.
    \item Run \(q\) synchronously working copies of AGS, but instead of performing Step 3, stop all
    the methods and select an interval with the highest characteristic among intervals from all the methods.
    \item If the interval with the highest characteristic belongs to problem i \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
    Other copies skip Step 3.
    \item To implement parallel processing, method picks \(p\) intervals at the previous two steps
    and performs \(p\) trials simultaneously.
  \end{itemize}
  The described method relies to \textbf{comparability of characteristics in different problems}.
\end{frame}

@@

\begin{frame}
  \frametitle{Index method of handling inequality constraints}
  \begin{enumerate}
    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  \begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
    the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
    Then solving all of the \(q\) problems by MIAGS will
    generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
    \begin{displaymath}
      \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
    \end{displaymath}
  \end{theorem}
\end{frame}

@@

\begin{frame}
  \frametitle{Generated test problems}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      Sets of test problems are obtained by applying the GCGen system to
      functions produced by generators of unconstrained test problems.
      The main features of the generated sets:
      \begin{itemize}
        \item \(q\)=100;
        \item Dimension 2, 3 or 4;
        \item Base fuctions GKLS or \(F_{GR}\), or a combination of them;
        \item Number of nonlinear constraints: 2.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
    \end{column}
  \end{columns}
  \footnotesize{GCGen is available here \textit{https://github.com/UNN-ITMM-Software/GCGen}}
\end{frame}

@@

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS \& \(F_{GR}\)-based \
        & 1 & 51434 & 90.20 & -    & - \\
        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
      \hline
      GKLS-based 2d \
        & 1 & 59066 & 97.53 & -    & - \\
        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS-based 3d \
        & 1 & 782544 & 1117.55 & -    & - \\
        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
      \hline
      GKLS-based 4d \
        & 1 & 14021720 & 15806.6 & -    & - \\
        & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
        & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
        & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Conclusions}
    \begin{itemize}
      \item implemented support of nonlinear functional constraints in the method
      for solving a set of global optimization problems;
      \item convergence of the implemented method was theoretically proven;
      \item conducted numerical experiments on sets, demonstrating uniform convergence (in the mentioned sense) and
      parallel speedup;
      \item demonstrated efficiency of the implemented method in case of solving a multi-objective problem.
    \end{itemize}
\end{frame}

@@

    sovrasov.vlad@gmail.com

>>>>

\begin{frame}
  \frametitle{Problem statement}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{displaymath}
        \begin{array}{cr}\\
          \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y:\\
           g_j(y)\leqslant 0, 1\leqslant j\leqslant m\},
        \end{array}
      \end{displaymath}
      where \(D\) is a hyperrectangle in \(\mathbb{R}^N\),
      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
      \begin{displaymath}
        \begin{array}{cr}\\
          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
        \end{array}
      \end{displaymath}
      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Problem statement}
  Such sets of problems can arise in the following cases:
  \begin{itemize}
    \item continuous global optimization problem with a discrete parameter, which takes a narrow range of values;
    \item a set of problems, obtained after scalarization of a multi-objective problem.
  \end{itemize}
\textbf{Possible approaches}:
  \begin{itemize}
    \item solve each problem independently;
    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
    prioritizing one of the problems.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Univariate Algorithm of Global Search}
  Sufficient convergence conditions:
  \begin{enumerate}
    \item \(G\ne\emptyset\), the source problem has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item During the optimization process estimations of \(L_i\) becomes large enough.
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Algorithm for solving a set of problems}
  The method relies on the previously introduced AGS and consists of the following steps:
  \begin{itemize}
    \item Create \(q\) copies of AGS.
    \item Run \(q\) synchronously working copies of AGS, but instead of performing Step 3, stop all
    the methods and select an interval with the highest characteristic among intervals from all the methods.
    \item If the interval with the highest characteristic belongs to problem \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
    Other copies skip Step 3.
    \item To implement parallel processing, method picks \(p\) intervals at the previous two steps
    and performs \(p\) trials simultaneously.
  \end{itemize}
  The described method relies to \textbf{comparability of characteristics in different problems}.
\end{frame}

@@

@@

\begin{frame}
  \frametitle{Generated test problems}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      Sets of test problems are obtained by applying the GCGen system to
      functions produced by generators of unconstrained test problems.
      The main features of the generated sets:
      \begin{itemize}
        \item \(q\)=100;
        \item Dimension 2, 3 or 4;
        \item Base fuctions GKLS or \(F_{GR}\), or a combination of them;
        \item Number of nonlinear constraints: 2.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
    \end{column}
  \end{columns}
  \footnotesize{GCGen is available here: \textit{https://github.com/UNN-ITMM-Software/GCGen}}
\end{frame}

@@

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS \& \(F_{GR}\)-based \
        & 1 & 51434 & 90.20 & -    & - \\
        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
      \hline
      GKLS-based 2d \
        & 1 & 59066 & 97.53 & -    & - \\
        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS-based 3d \
        & 1 & 782544 & 1117.55 & -    & - \\
        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
      \hline
      GKLS-based 4d \
        & 1 & 14021720 & 15806.6 & -    & - \\
        & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
        & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
        & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Conclusions}
    \begin{itemize}
      \item Implemented support of nonlinear functional constraints in the method
      for solving a set of global optimization problems;
      \item convergence of the implemented method was theoretically proven;
      \item conducted numerical experiments on sets, demonstrating uniform convergence (in the mentioned sense) and
      parallel speedup;
      \item demonstrated efficiency of the implemented method in case of solving a multi-objective problem.
    \end{itemize}
\end{frame}

@@

    sovrasov.vladislav@itmm.unn.ru
